Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-04-20 appendix.qmd
Added cluster colors

tsne_pbmc_scaled_with_cluster <- tsne_pbmc_scaled |>
  mutate(cluster = if_else((emb1 <= 0.4) & (emb2 <= 0.6), "cluster1", if_else((emb1 >= 0.75) & (emb2 <= 0.25), "cluster3", "cluster2"))) |>
  mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2", "cluster3")))

@@

trimesh_removed_pbmc_best <- ggplot() + 
  geom_point(
    data = tsne_pbmc_scaled_with_cluster,
    aes(
      x = emb1,
      y = emb2,
      color = cluster
    ),
    alpha = 0.2,
    size = 0.1
  )  +
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
  interior_annotation("a1", cex = 2) +
  theme(
    aspect.ratio = 1
  )
```

@@

pbmc_proj_tsne_model1_best <- plot_proj(
  proj_obj = proj_obj1, 
  point_param = c(1.5, 0.2), # size, alpha, color
  plot_limits = c(-0.65, 0.65), 
  title = "a2", 
  cex = 2, 
  axis_text_size = 4,
  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))

@@

pbmc_proj_tsne_model2_best <- plot_proj(
  proj_obj = proj_obj2, 
  point_param = c(1.5, 0.2), # size, alpha, color
  plot_limits = c(-1, 0.75), 
  title = "a3", 
  cex = 2, 
  axis_text_size = 4,
  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))

>>>>

distance_df_small_edges_pbmc <- distance_pbmc |>
  filter(distance < 0.1)#benchmark_pbmc

tsne_pbmc_scaled_with_cluster <- tsne_pbmc_scaled |>
  mutate(cluster = if_else((emb2 <= -emb1 + 0.8) & (emb1 < 0.5), "cluster1", if_else(emb2 >= -2 * emb1 + 1.79, "cluster3", "cluster2"))) |>
  mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2", "cluster3")))

@@

trimesh_removed_pbmc_best <- ggplot() + 
  geom_point(
    data = tsne_pbmc_scaled_with_cluster,
    aes(
      x = emb1,
      y = emb2,
      color = cluster
    ),
    alpha = 0.2,
    size = 0.1
  )  +
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
  interior_annotation("b1", cex = 2) +
  theme(
    aspect.ratio = 1
  )
```

@@

pbmc_proj_tsne_model1_best <- plot_proj(
  proj_obj = proj_obj1, 
  point_param = c(1.5, 0.2), # size, alpha, color
  plot_limits = c(-0.65, 0.65), 
  title = "b2", 
  cex = 2, 
  axis_text_size = 4,
  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))

@@

pbmc_proj_tsne_model2_best <- plot_proj(
  proj_obj = proj_obj2, 
  point_param = c(1.5, 0.2), # size, alpha, color
  plot_limits = c(-1, 0.75), 
  title = "b3", 
  cex = 2, 
  axis_text_size = 4,
  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
